Identifying regime transitions for water governance at a basin scale

### 2022-04-28 paper/main/01_introduction.tex
fix(introducion,-conclusion,-meta): refactor introduction
according to ShuaiWang and XutongWu comments

Governance is essentially about ``who gets what, when and how''. Specifically, water governance refers to the political, social, economic, and administrative systems that influence the use and management of water %! citation.
For water resources in populated areas, missing governance means missing sustainability, and a first important step in understanding transitions toward successful water governance is identifying the different regimes.
\cite{undpwatergovernancefacilityWaterGovernanceIssue}.
Corresponding to ``who gets what, when and how'', the United Nations Development Programme (UNDP) suggests three key dimensions of water use are decided by the water governance directly: ``When and what water to use?'' (stress), ``How water provides different services to well-beings?'' (purpose), and ``Who can use water equally and efficiently?'' (allocation)
\cite{undpwatergovernancefacilityWaterGovernanceIssue, mariajacobsonUserguideassessing2013, kjellenWatergovernanceperspective2015}.

As a temporary stable state in structure, function, and dominant controls, a specific regime of water governance maintains by concreted intertwines within human-water systems (such as management, institutions, and exploitations)
\cite{carpenterEarlyWarningsRegime2011}.
However, large and persistent changes in key properties may lead to a loss of systematical stability, potentially resulting in a regime shift with impacts on system outcomes and widespread cascading effects
\cite{rochaCascadingregimeshifts2018, gregrCascadingsocialecologicalcosts2020}.
Therefore, as both signals and consequences of substantive changes in human-water systems, regime shifts align with changing three key dimensions of water governance (stress, purpose, and allocation), sometimes leading to new challenges in sustainability %! UNDP year
\cite{undpwatergovernancefacilityWaterGovernanceIssue, mariajacobsonUserguideassessing2013, kjellenWatergovernanceperspective2015}.
First, water stress depends not only on climate (with increasing scarcity and uncertainty in many regions) but also on the increasingly insatiable demands from economic activities such as irrigation and industry; water storage can resolve some but not all of these issues
\cite{greveGlobalassessmentwater2018,wadaHumanwaterinterface2017,qinFlexibilityintensityglobal2019}.
Second, the purpose of how water services human well-being is to consider trade-offs between consumptive uses (e.g., drinking and food production) and non-consumptive uses (e.g., energy production)
\cite{liuWaterscarcityassessments2017,florkeWatercompetitioncities2018,kleemannQuantifyinginterregionalflows2020}.
Third, the allocation of water across the whole basin is influenced not only by regionally socio-economic and environmental context but also by systematically regulating resources
\cite{roobavannanRoleSectoralTransformation2017,speedBasinwaterallocation2013}.
Despite regime shifts in water governance related to substantive changes in any of the three dimensions, separately considering their intertwines within human-water systems can lead to holistic failure in governing water.
The lack of a comprehensive but straightforward approach to identifying changes in water governance regimes challenges sustainability, and filling this gap can well align human and water systems (Figure~\ref{fig:framework}).

As an informative example, we focus on the Yellow River Basin (YRB, see \textit{Appendix} Methods S1 and Figure S1 for details), the fifth-largest river in the world.
With drastically anthropogenic intervention, the YRB experienced the most intense water governance challenges in China, leading to long-standing sustainability barriers.
Since the 1960s, the implementation of conservation measures, regulation reservoirs, and levee constructions have contained the issues caused by high-sediment loads
\cite{wangReducedsedimenttransport2016,wuEvolutioneffectssocialecological2020}.
However, water over-use has led to depletions of the over-burdened river, creating new challenges and new governance practices, including water use regulation and water transfer across basins
\cite{xiaDevelopmentWaterAllocation2012}.
Today, it is still impossible to completely solve water stress, trade-offs between ecosystem services, or lopsided development in different regions in the YRB; -``who gets water, when and how‘’ is still an open question for sustainable water governance
\cite{wangYellowRiverwater2019,wohlfartSocialecologicalchallengesYellow2016}.
However, confronting governance challenges induced by environmental, economic, social, and political factors, numerous governance practices have led the YRB to be among the most drastically-governing large river basins worldwide
Identifying regime shifts in water governance within the YRB can thus provide crucial insights into rapidly-changing big river basins and how governance may respond to meeting challenges to their sustainability.

>>>>

Water governance refers to the political, social, economic, and administrative systems that influence the use and management of water %! citation.
For populated large river basins, missing governance means missing sustainability, and a first important step in understanding the transitions with successful water governance is identifying the different regimes.
\cite{undpwatergovernancefacilityWaterGovernanceIssue}.
Regime of water governance maintains by concreted intertwines within human-water systems (such as management, institutions, and exploitations), as a stable state in structures and functions
\cite{carpenterEarlyWarningsRegime2011,rochaCascadingregimeshifts2018, gregrCascadingsocialecologicalcosts2020}.
Therefore, as both signals and consequences of substantive changes in human-water systems, regime shifts sometimes leading to new challenges in water governance.
The lack of a comprehensive but straightforward approach to identifying changes in water governance regimes challenges sustainability, and filling this gap can well align human and water systems (Figure~\ref{fig:framework}).

Governance is essentially about ``who gets what, when and how'', the United Nations Development Programme (UNDP) thus suggested three key dimensions of water use are decided by the water governance directly: ``When and what water to use?'' (stress), ``How water provides different services to well-beings?'' (purpose), and ``Who can use water equally and efficiently?'' (allocation)
\cite{undpwatergovernancefacilityWaterGovernanceIssue, mariajacobsonUserguideassessing2013, kjellenWatergovernanceperspective2015}. %! UNDP year
First, water stress depends not only on climate (with increasing scarcity and uncertainty in many regions) but also on the increasingly insatiable demands from economic activities such as irrigation and industry; water storage can resolve some but not all of these issues
\cite{greveGlobalassessmentwater2018,wadaHumanwaterinterface2017,qinFlexibilityintensityglobal2019}.
Second, the purpose of how water services human well-being is to consider trade-offs between consumptive uses (e.g., drinking and food production) and non-consumptive uses (e.g., energy production)
\cite{liuWaterscarcityassessments2017,florkeWatercompetitioncities2018,kleemannQuantifyinginterregionalflows2020}.
Third, the allocation of water across the whole basin is not only decided by regionally socio-economic and environmental context but also influenced by systematically regulating
\cite{roobavannanRoleSectoralTransformation2017,speedBasinwaterallocation2013}.
Despite regime shifts in water governance related to substantive changes in any of the three dimensions, separately considering their intertwines within human-water systems can lead to holistic failure in governing water.

The Yellow River Basin (YRB), fifth-largest river in the world, was most in need of integrated water governance because drastically anthropogenic intervention led to intense governance challenges in sustainability (see \textit{Appendix} Methods S1 and Figure S1 for details).
Since the 1960s, the implementation of conservation measures, regulation reservoirs, and levee constructions have contained the governance issues troubled for thousands years of high sediment loads
\cite{wangReducedsedimenttransport2016,wuEvolutioneffectssocialecological2020}.
However, decreased streamflows and water over-use then led to depletions of the over-burdened river, creating new challenges and new governance practices, including water use regulation and water transfer across basins
\cite{xiaDevelopmentWaterAllocation2012}.
Today, it is still impossible to completely solve water stress, trade-offs between ecosystem services, or lopsided development in different regions in the YRB; -``who gets water, when and how‘’ is always an open question for sustainable development
\cite{wangYellowRiverwater2019,wohlfartSocialecologicalchallengesYellow2016}.
Confronting governance challenges induced by environmental, economic, social, and political factors, numerous governance practices have led the YRB to be among the most drastically-governing large river basins worldwide
Identifying regime shifts in water governance within the YRB can thus provide crucial insights into rapidly-changing big river basins and how governance may respond to meeting challenges to their sustainability.

### 2022-04-28 paper/main/05_conclusion.tex
fix(introducion,-conclusion,-meta): refactor introduction
according to ShuaiWang and XutongWu comments

Three dimensions of water governance changes along with the hydrosocial cycle transition: water stress, services purpose, and water allocation, affecting ``who gets water, when and how''. By integrating them, we developed an Integrated Water Governance Index (IWGI) to detect regime shifts in water governance. Applying the index to a rapidly-changing large river basin (the Yellow River Basin, China) describes how water governance shifts between three regimes over half a century (massive supply regime; governance transformation regime; and adaptation oriented regime, respectively). Our approach quantitatively identifies the genearl schema for water governance regimes in the YRB, in line with previous theoretical analysis with a representative transition process. Linking regime shifts to the underlying causes, the implementation of IWGI offers a comprehensive and straightforward way to interpret changes in intertwines of water governance, hydrosocial transition, and human-water relationships.

>>>>

Three dimensions of water governance change along with the hydrosocial cycle transition: water stress, services purpose, and water allocation, affecting ``who gets water, when and how''. We developed an Integrated Water Governance Index (IWGI) to detect regime shifts in water governance by integrating them. Applying the index to a rapidly-changing large river basin (the Yellow River Basin, China) describes how water governance shifts between three regimes over half a century (massive supply regime; governance transformation regime; and adaptation oriented regime, respectively). Our approach quantitatively identifies the general schema for water governance regimes in the YRB, in line with previous theoretical analysis with a representative transition process. Linking regime shifts to the underlying causes, the implementation of IWGI offers a comprehensive and straightforward way to interpret changes in intertwines of water governance, hydrosocial transition, and human-water relationships.

### 2022-04-28 paper/main/manuscript.tex
fix(introducion,-conclusion,-meta): refactor introduction
according to ShuaiWang and XutongWu comments

\abstract{
    In many large river basins, water governance regimes from natural to social-ecological or `hydrosocial' situations profoundly change who gets water, when, and how. Therefore, identifying changes in governance regimes is critical to understanding the transition and guiding the efficient and sustainable use of water. We combined three main dimensions of water governance (supply, purpose and allocation) to develop a quantitative Integrated Water Governance Index (IWGI) that can detect regime shifts in water governance at a basin scale. Applying the index to a rapidly-changing large river basin (the Yellow River Basin, China) describes how water governance shifts between three regimes over half a century (massive supply regime; governance transformation regime; and adaptation oriented regime, respectively). Our application of the IWGI offers a comprehensive and straightforward way to link the shifts in water governance to environmental, economic, social and political factors. Moreover, as a potentially widespread transition schema for water governance regimes, the different governance challenges that occurred at different phases of the transition of the Yellow River also suggest potentially helpful guidelines on the governance of big river basins in general.
}

\keywords{
    Regime shifts,
    Water governance,
    Transformation,
    Sustainability
}

>>>>

\abstract{
    In many large river basins, water governance regimes from natural to social-ecological or `hydrosocial' situations profoundly change who gets water, when, and how.
    Identifying changes in governance regimes is critical to understanding the transition and guiding the efficient and sustainable use of water.
    We combined three main dimensions of water governance (supply, purpose and allocation) to develop a quantitative Integrated Water Governance Index (IWGI) to detect regime shifts in water governance at a basin scale.
    Applying the index to a rapidly-changing large river basin (the Yellow River Basin, China) describes how water governance shifts between three regimes over half a century. (massive supply regime; governance transformation regime; and adaptation oriented regime, respectively).
    In the YRB, the regime shifts' underlying causes are increasing water supply and demand through the construction of reservoirs and expansion of irrigated areas before the governance transformation, while are re-allocation and regulation after the transformation.
    Our application of the IWGI offers a comprehensive and straightforward way to link the water governance regimes to the hydrosocial transition, following challenges in sustainability.
}

\keywords{
    Regime shifts,
    Water governance,
    Transformation,
    Social-hydrology,
    Sustainability
}
